An Adolescent Mini-Neptune in the Kepler Field

### 2021-07-12 paper/ms.tex
update ms

\defcitealias{cantatgaudin_gaia_2018}{CG18}
\defcitealias{kounkel_untangling_2019}{KC19}
\defcitealias{meingast_2021}{M21}

% Cluster Difference Imaging Photometric Survey. III.
% Rotation and Lithium in an Open Cluster Spanning 500\,Parsecs
\title{
  An Adolescent Mini-Neptune in the Kepler Field
}

@@

\begin{abstract}
  Kepler\,1627A is a G8V star previously found to host a
  $3.6\,R_\oplus$ mini-Neptune on a 7.2\,day orbit.  The star was
  selected for observation by the Kepler satellite because it is
  nearby ($d\approx 333\,{\rm pc}$) and it resembles the Sun.  Combining
  data from Gaia with rotation periods from TESS, we show that
  Kepler\,1627 is a member of the $35\pm XX$\,Myr old $\delta$~Lyr
  cluster.  Kepler\,1627Ab is therefore the youngest planet with a
  precise age found to date by the main Kepler mission.
  Analyzing the photometry, we find evidence that starspot crossings
  distort the average transit shape, and induce transit timing
  variations that suggest the planet is on a prograde orbit.  The 
  flare arrival times may show hints of being
  non-Poissonian, but the statistical evidence is not overwhelming ($<3$-$\sigma$).  In the future,
  spectroscopic studies of \pn\ may help clarify the evolution of this
  mini-Neptune's orbit and atmosphere.  More broadly, the
  combination of Gaia and TESS data offers an opportunity to
  expand the census of age-dated planets; we provide a literature
  concatenation of young stars to support this effort.
\end{abstract}

@@

For instance, the ``mini-Neptune'' planets, thought to be made of
molten rocky cores and extended atmospheric envelopes of hydrogen and
helium, are expected to shrink in size by factors of several over
their first $10^8$ years
\citep{owen_constraining_2020,kite_atmosphere_2020}.  Specifically,
they start with sizes of 4--12\,$R_\oplus$ at the time of disk
dispersal ($\lesssim$$10^7$\,years), and shrink to sizes of
2--4\,$R_\oplus$ by 10$^8$ years.  This change is driven by a
combination of stellar irradiation and internal heat that powers an
outflow, which can eventually deplete or entirely strip the envelope
from the rocky core \citep{Owen_Wu_2013,gupta_sculpting_2019}.
Discovering young planets and measuring their compositions and
atmospheric outflows are key steps toward testing this paradigm, which
is often invoked to explain the observed radius distribution of both
mature exoplanets \citep{Fulton_et_al_2017}, as well as younger
planets \citep{bouma_cluster_2020}.

@@

At the time of the main Kepler mission (2009--2013), only four open
clusters were known in the Kepler field: NGC\,6866, NGC\,6811,
NGC\,6819, and NGC\,6791, with ages spanning 0.7\,Gyr to 9\,Gyr
\citep{meibom_kepler_2011}.  Since that time, analyses of the
kinematic, photometric, and astrometric Gaia data have expanded our
knowledge of open cluster and moving group memberships \citep[{\it
e.g.},][]{cantatgaudin_gaia_2018,zari_3d_2018,kounkel_untangling_2019,Meingast2021}.
As part of our Cluster Difference Imaging Photometric Survey (CDIPS,
\citealt{bouma_cdipsI_2019}), we concatenated the available analyses
from the literature, which yielded a list of candidate young and
age-dated stars (see Appendix~\ref{app:targetlist}).

Comparing our young star list against the Kepler field yielded two
discoveries.  The first, to be discussed in an upcoming analysis by J.
Curtis et al{.}, is that Kepler observed the $\approx$350\,Myr open
cluster Theia\,520 (UBC\,1).  At least six Kepler planets are known in
the cluster, from the Kepler-52 and Kepler-968 systems
\citep{rowe_validation_2014,jontof-hutter_following_2021}.  The second
discovery, and the focus of this work, is that Kepler observed a
portion of the $\delta$\,Lyr cluster, shown in
Figure~\ref{fig:skychart}.

@@

Could the transit signal be produced by anything other than a planet
orbiting this near-solar analog?  \citet{morton_false_2016} considered
the false positive scenarios of a foreground eclipsing binary, a
hierarchical triple-star binary system, and a background eclipsing
binary.  Based on the transit shape, \citet{morton_false_2016} found
the latter scenario to be the only possible case, with a
(model-dependent) probability of $\approx10^{-5}$.  While this
statistical validation is reassuring, ``validated'' planets have been
demonstrated to be false positives \citep{shporer_three_2017}.  Most
failures of the statistical validation approach arise when attempting
to separate the smallest stars from identically-sized gas giant
planets.  With a size of $\approx 3.6\,R_\oplus$, Kepler\,1627b does
not meet this concern.  The fact that we detect a low-mass stellar
companion in archival Keck-NIRC2 Kp-band (2.12\,$\mu $m) imaging does
not change the conclusion: the companion contributes 1\% of the flux
observed by Kepler, which combined with the observed transit shape
cannot be invoked as an explanation for the observed data
(Appendix~\ref{app:companionstar}).

{\bf todo: LGB to check numbers on full-dataset transit ecc and b} 
Two additional new lines of evidence support the planetary
interpretation.  First, the $\approx$100 days of 1-minute
short-cadence data reduces the cadence-induced smearing enables a more
precise measurement of the transit impact parameter
\citep{kipping_binning_2010}.  The planet transits with $b<0.YY$
(3-$\sigma$) from the 1-minute cadenced data, compared to $<0.XX$
(3-$\sigma$) from the 30-minute data.

{\bf todo: Kento to check numbers on per-transit analysis} 
The second line of evidence came from our transit timing analysis. We
isolated each of the 144 observed transits to within $\pm4.5$\,hr of
each transit, and fitted each window with both {\it i)} a local
second-order polynomial and transit, and {\it ii)} a local linear
trend.  We let the mid-time of each transit float, and then calculated
the residual between the measured mid-time and that of a perfectly
periodic orbit.  This residual, the ``transit timing variation''
(TTV), is plotted in Figure~\ref{fig:ttvslope} against the local
linear slope.  Only a handful of Kepler Objects of Interest have shown
this correlation \citep{holczer_time_2015}, which is most readily
interpreted as a TTV induced by unresolved starspot crossings
\citep{mazeh_time_2015}.  This is of course only possible if the
planet is transiting the primary star, which excludes the any of the
aforementioned false positive scenarios.  It also implies that the
planet's orbit is prograde.  The latter interpretation assumes that
the dominant photometric variability is induced by dark spots, rather
than bright faculae.  Given the observed transition of Sun-like
stellar variability from spot to faculae-dominated regimes, we expect
this inference to be reasonably secure
\citep{shapiro_are_2016,montet_long-term_2017,reinhold_stellar_2020}.

>>>>

\title{
  An Adolescent Mini-Neptune in the Kepler Field
}

@@

\begin{abstract}
  Kepler\,1627A is a G8V star previously found to host a
  $3.6\,R_\oplus$ mini-Neptune on a 7.2\,day orbit.  The star was
  selected for observation by the Kepler satellite because it is
  nearby ($d\approx 333\,{\rm pc}$) and it resembles the Sun.  Combining
  data from Gaia with rotation periods from TESS, we show that
  Kepler\,1627 is a member of the $35\pm XX$\,Myr old $\delta$~Lyr
  cluster.  Kepler\,1627Ab is therefore the youngest planet with a
  precise age found to date by the main Kepler mission.
  Analyzing the photometry, we find evidence that starspot crossings
  distort the average transit shape, and induce transit timing
  variations that suggest the planet is on a prograde orbit.  The 
  flare arrival time distribution is most likely Poissonian, though
  there may be hints ($<3$-$\sigma$) of additional flares at the
  orbital and synodic periods.  In the future, spectroscopic studies
  of \pn\ may help clarify the evolution of this mini-Neptune's orbit
  and atmosphere.  More broadly, the combination of Gaia and TESS data
  offers an opportunity to expand the census of age-dated planets; we
  provide a literature concatenation of young stars to support this
  effort.
\end{abstract}

@@

For instance, the ``mini-Neptune'' planets, thought to be made of
molten rocky cores and extended atmospheric envelopes of hydrogen and
helium, are expected to shrink in size by factors of several over
their first $10^8$ years
\citep{owen_constraining_2020,kite_atmosphere_2020}.  Specifically,
they start with sizes of 4--12\,$R_\oplus$ at the time of disk
dispersal ($\lesssim$$10^7$\,years), and shrink to sizes of
2--4\,$R_\oplus$ by 10$^8$ years.  This change is driven by a
combination of stellar irradiation and internal heat that powers an
outflow, which can eventually deplete or entirely strip the envelope
from the rocky core \citep{Owen_Wu_2013,gupta_sculpting_2019}.
Discovering young planets and measuring their compositions and
atmospheric outflows are key steps toward testing this paradigm, which
is often invoked to explain the observed radius distribution of 
mature exoplanets \citep{Fulton_et_al_2017}.

@@

At the time of the main Kepler mission (2009--2013), only four open
clusters were known in the Kepler field: NGC\,6866, NGC\,6811,
NGC\,6819, and NGC\,6791, with ages spanning 0.7\,Gyr to 9\,Gyr
\citep{meibom_kepler_2011}.  Since that time, analyses of the
kinematic, photometric, and astrometric Gaia data have expanded our
knowledge of open cluster and moving group memberships \citep[{\it
e.g.},][]{cantatgaudin_gaia_2018,zari_3d_2018,kounkel_untangling_2019,Meingast2021,kerr_stars_2021}.
As part of our Cluster Difference Imaging Photometric Survey (CDIPS,
\citealt{bouma_cdipsI_2019}), we concatenated the available analyses
from the literature, which yielded a list of candidate young and
age-dated stars (see Appendix~\ref{app:targetlist}).

Comparing our young star list against the Kepler field yielded two
discoveries.  The first, to be discussed in an upcoming analysis by J.
Curtis et al{.}, is that Kepler observed the $\approx$350\,Myr open
cluster Theia\,520 (UBC\,1).  At least six Kepler planets are known in
the cluster from the Kepler-52 and Kepler-968 systems
\citep{rowe_validation_2014,jontof-hutter_following_2021}.  The second
discovery, and the focus of this work, is that Kepler observed a
portion of the $\delta$\,Lyr cluster \citep{stephenson_possible_1959,eggen_photometric_1968}.
The overlap between the Kepler field and the cluster is shown in
Figure~\ref{fig:skychart}.

@@

Could the transit signal be produced by anything other than a planet
orbiting this near-solar analog?  \citet{morton_false_2016} considered
the false positive scenarios of a foreground eclipsing binary, a
hierarchical triple-star binary system, and a background eclipsing
binary.  Based on the transit shape, \citet{morton_false_2016} found
the latter scenario to be the only possible case, with a
(model-dependent) probability of $\approx10^{-5}$.  While this
statistical validation is reassuring, ``validated'' planets have been
shown to be false positives \citep{shporer_three_2017}.  Most
failures of the statistical validation approach arise when attempting
to separate the smallest stars from identically-sized gas giant
planets.  With a size of $\approx 3.6\,R_\oplus$, Kepler\,1627b does
not meet this concern.  The fact that we detect a low-mass stellar
companion in archival Keck-NIRC2 Kp-band (2.12\,$\mu $m) imaging does
not change the conclusion: the companion star contributes 1\% of the
flux observed by Kepler.  Combined with the observed transit depth and
shape (Appendix~\ref{app:companionstar}), a planet transiting the
primary star is the only plausible interpretation of the
data.

{\bf todo: LGB to check numbers on full-dataset transit ecc and b} 
Two additional new lines of evidence confirm the planetary
interpretation.  First, the $\approx$100 days of 1-minute
cadenced Kepler observations reduce the cadence-induced smearing,
enabling a more precise measurement of the transit impact parameter
\citep{kipping_binning_2010}.  The planet transits with $b<0.YY$
(3-$\sigma$) from the 1-minute cadenced data, compared to $<0.XX$
(3-$\sigma$) from the 30-minute data.

{\bf todo: Kento to check numbers on per-transit analysis} 
The second line of evidence comes from our transit timing analysis. We
isolated each of the 144 observed transits to within $\pm4.5$\,hr of
each transit, and fitted each window with both {\it i)} a local
second-order polynomial and transit, and {\it ii)} a local linear
trend.  We let the mid-time of each transit float, and then calculated
the residual between the measured mid-time and that of a perfectly
periodic orbit.  This residual, the ``transit timing variation''
(TTV), is plotted in Figure~\ref{fig:ttvslope} against the local
linear slope.  Only a handful of Kepler Objects of Interest have shown
this correlation \citep{holczer_time_2015}, which is most readily
interpreted as a TTV induced by unresolved starspot crossings
\citep{mazeh_time_2015}.  This is of course only possible if the
planet is transiting the primary star, which excludes the any of the
aforementioned false positive scenarios.  It also implies that the
planet's orbit is prograde.  The latter interpretation assumes that
the dominant photometric variability is induced by dark spots, rather
than bright faculae.  Given the observed transition of Sun-like
stellar variability from spot to faculae-dominated regimes, we expect
this inference to be reasonably secure
\citep{shapiro_are_2016,montet_long-term_2017,reinhold_stellar_2020}.
